---
title: "Data & Exploration"
output: 
  html_document:
    toc: FALSE
---
# Data and Data Exploration:

The following excerpt are the processing and data exploration steps completed for the `Influenza Analysis`.

<br>

# Data Wrangling

```{r}
#load needed packages. make sure they are installed.
library(readxl) #for loading Excel files
library(dplyr) #for data processing
library(here) #to set paths
```


## Loading Raw Data

```{r}
#path to data

data_location <- here::here("data","raw_data","SympAct_Any_Pos.Rds")

#load data. 

rawdata <- readRDS(data_location)

#take a look at the data
dplyr::glimpse(rawdata)
str(rawdata)
summary(rawdata)




```


## Processing Data

```{r}
#remove variable Score, Total, FluA, FluB, Dxname, or activity in name
#chose to use direct pipes since the actions were so similar and I didn't wnat to repeat myself too much in the comments
##select(-contains("")) == removes columns with the whatever is inside the "quotes".
rawdata1<-rawdata%>%
  select(-contains("Score"))%>% #removed cols containing Score
  select(-contains("Total"))%>% #removed cols containing Total
  select(-contains("FluA"))%>% #removed cols containing FluA
  select(-contains("FluB"))%>% #removed cols containing FluB
  select(-contains("DxName"))%>% #removed cols containing DxName
  select(-contains("Activity"))%>% #removed cols containing Activity
  select(-c(Unique.Visit)) # -c removes exactly named cols, remove col Unique.Visit
  
#take a look at  cleaned data  removed cols directed by assessment page
glimpse(rawdata1)  
str(rawdata1)
summary(rawdata1)

#the data set is not quite like the assignment I am 5 over in observations
#noted that BodyTemp has 5 NA's, need to remove

rawdata2<-rawdata1%>%
  na.omit() # omit NAs in the data set

#take a look at  cleaned data to assignment specs (760 obv. of 32 variables)
glimpse(rawdata2)  
str(rawdata2)
summary(rawdata2)
```


## Saving Data

```{r}
# location to save file
save_data_location <- here::here("data","processed_data","processeddata.rds")

saveRDS(rawdata2, file = save_data_location)
```


<br>
# Data Exploration

---
First loading all the default settings and preliminary programs.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for data processing
library(here) #to set paths
```

Path to Processed Data and loading
```{r}
data_location <- here::here("data","processed_data","processeddata.rds")
data<- readRDS(data_location)
```
View Summary of new Data, subsetted from the original raw data.
```{r}
str(data) #check structure of R object
glimpse(data) #details of the data up to 4 levels, empty values = "" 
summary(data)
```

Important variables:  Outcome= Body temperature and our main categorical outcome is Nausea. 
We want to see if the other symptoms are correlated with (predict) those outcomes.

Limited selection of other symptoms (Selection of 5, based on vaugely off of the PeptoBismol jingle ):  

1. Nausea
2. CoughYN
3. Pharyngitis
4. Vomit
5. Diarrhea

Reduce data to limited variables and BodyTemp. Preview to confirm
```{r}
SelectData<-data%>% #make into a new data frame
  #select limited variables and BodyTemp
  select(Nausea, 
         CoughYN,
         Pharyngitis,
         Vomit,
         Diarrhea,
         BodyTemp)%>%
  #filter for only positive Nausea data, per instruction of prompt as it being the main categorical outcome
  filter(Nausea=="Yes")

summary(SelectData)
```

##Summary Table
```{r}
SelectData%>%
  #grouping by Nausea, the main categorical outcome
  group_by(Nausea,
           CoughYN, 
           Pharyngitis, 
           Vomit, 
           Diarrhea)%>%  
  #summarise the data, get observances and mean temp, calculate the standard deviation
  summarise(
    n=n(), #to count the number of observances
    mean=mean(BodyTemp),#calculate mean body temp
    sd=sd(BodyTemp) #calculate sd (Standard Deviation) of Body temp
  )%>%
  #arrange by highest number of n
  arrange(desc(n)) %>%
  print()

```
Most cases with observed experience Nausea, Cough, Pharyngitis with a mean Body temperature of 99.1.
Only once case where it was just Nausea and none of the other selected variables.

Save of temp and positive for nausea
```{r}
save_data_location <- here::here("data","processed_data","SelectData.rds")

saveRDS(SelectData, file = save_data_location)
```


Not satisfied with my version of summary table... so looking at other examples to see how to do this better.
Found someone using this table1 program. Let's see how it goes.
```{r}

library(table1)# saw this used in someone else's code, trying it out
table1(~   BodyTemp+
           CoughYN+
           Pharyngitis+ 
           Vomit+ 
         # the | splits it to be factor's col variables-- this case of Nausea-.
           Diarrhea | Nausea, 
       #use prior data since selecteddata only shows the positive nausea hits
       data=data, 
       overall="Total")

```
This is a very nice table, Need to look into about putting in BodyTemp as the factors in this and not just Nausea


```{r}

```

##Histogram/Density plot

Code here is copy pasted and changes in the data source and fill aspects are based on the different comparisons.

By Nausea
```{r}
data%>% #using original data to show Yes and No for Nausea
  ggplot(aes(x=BodyTemp, fill=Nausea)) +
    geom_density(adjust=1.5,
                 alpha=.5)+
  labs(title= "Body temperature of people experencing nausea")
```
By Vomit
```{r}
SelectData%>%
  ggplot(aes(x=BodyTemp, fill=CoughYN)) +
    geom_density(adjust=1.5,
                 alpha=.5)+
    labs(title= "Body temperature of people experiencing nausea with a cough")

```
By Cough
```{r}
SelectData%>%
  ggplot(aes(x=BodyTemp, fill=Pharyngitis)) +
    geom_density(adjust=1.5,
                 alpha=.5)+
    labs(title= "Body temperature of people experiencing nausea with pharyngitis")

```

```{r}
SelectData%>%
  ggplot(aes(x=BodyTemp, fill=Vomit)) +
    geom_density(adjust=1.5,
                 alpha=.5)+
    labs(title= "Body temperature of people experiencing nausea and vomitted")

```

```{r}
SelectData%>%
  ggplot(aes(x=BodyTemp, fill=Diarrhea)) +
    geom_density(adjust=1.5,
                 alpha=.5)+
    labs(title= "Body temperature of people experiencing nausea with diarrhea")

```

##Scatterplots and Boxplots

```{r}
#Nausea positive

SelectData%>% #only positive with Nausea
  ggplot(aes(y=BodyTemp))+
  geom_violin(aes(x=Pharyngitis, color="Pharyngitis", alpha=.5))+
  geom_violin(aes(x=Vomit, color="Vomit", alpha=.5))+
  geom_violin(aes(x=Diarrhea, color="Diarrhea", alpha=.5))+
  geom_jitter(aes(x=Pharyngitis, color="Pharyngitis"))+
  geom_jitter(aes(x=Vomit, color="Vomit"))+
  geom_jitter(aes(x=Diarrhea, color="Diarrhea"))+
  labs(title="Body temperature Violin graph of symptoms that occured with Nausea")+
  xlab("Symptom experienced with Nausea")

```

Need to reshape data to get the right output in ggplot
I think I need to rotate the table to long form? 
```{r}
#Pivot Select Data
PSData<-SelectData%>%
  group_by(CoughYN, 
           Pharyngitis, 
           Vomit, 
           Diarrhea)%>%
  summarize( 
    Temp=mean(BodyTemp),#calculate mean body temp
    sd=sd(BodyTemp))%>%
#really unsure if this was correct
  pivot_longer(
    cols = c(CoughYN, Pharyngitis,Vomit,Diarrhea),
    names_to = "Symptoms",
    values_to = "Presence" )

glimpse(PSData)
```
Save Pivoted Data of mean temp and positive for nausea
```{r}
save_data_location <- here::here("data","processed_data","PivotSelect.rds")

saveRDS(PSData, file = save_data_location)
```


```{r}
PSData%>%
  ggplot(aes(x=Symptoms, y=Temp))+
  geom_boxplot(aes(color=Presence))+
  labs(title="Body temperature boxplot of symptoms that occured with Nausea")+
  xlab("Symptom experienced with Nausea")

```
