---
title: "Machine Learning Results (Mod11)"
output: 
  html_document:
    toc: FALSE
---

# Machine Learning Results
Loading all the default settings and preliminary programs.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for data processing and all dplyr related programs
library(here) #to set paths
library(tidymodels)      # for the recipes package, along with the rest of tidymodels
library(skimr)           # for variable summaries

```

Path to Processed Data and loading of cleaned data

```{r}
data_location <- here::here("data","processed_data","processeddata.rds")
data<- readRDS(data_location)
```

**Reminder**: Outcome of interest is **Body Temp**; Categorical outcome is **Nausea**; Predictor= **RunnyNose**
```{r}
glimpse(data)
```
Feature / Variable removal

* Remove severity score data
```{r}
OrderedData<-data%>%
  #remove YN variables for those variables with severity factors
  select(-WeaknessYN,-CoughYN,-MyalgiaYN,-CoughYN2)%>%
  #code symptom severity factors as ordinal
  mutate(Weakness=as.ordered(Weakness), CoughIntensity=as.ordered(CoughIntensity), Myalgia=as.ordered(Myalgia))%>%
  #Order severity in ordered factors: None<mild<Moderate<Severee
 mutate_at(vars(Weakness, CoughIntensity, Myalgia),
           list(~factor(.,levels = c("None","Mild","Moderate","Severe"),ordered = TRUE)))

summary(OrderedData)
```
Low variance predictors
view summary of Ordered Data <50 entries
```{r}
Mod11Data<-OrderedData%>%
  select(-Hearing, -Vision)

glimpse(Mod11Data)  
```

location to save file
```{r}
save_data_location <- here::here("data","processed_data","Mod11Data.rds")

saveRDS(Mod11Data, file = save_data_location)
```
# Module 11 Contributor: Monica Chan

## Body Temperature vs. all predictors

Below I am creating the new full model recipe for body temperature, making a path to the linear modeling, and combining them into a workflow to create a fitted object.
