---
title: "Machine Learning Results (Mod11)"
output: 
  html_document:
    toc: FALSE
---

---
title: "Analysis Script Module 9"
author: "MYC"
date: "10/20/2021"
output: 
  html_document:
    toc: FALSE
---

Loading all the default settings and preliminary programs.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for data processing and all dplyr related programs
library(here) #to set paths
library(tidymodels)      # for the recipes package, along with the rest of tidymodels
library(skimr)           # for variable summaries

```

Path to Processed Data and loading of cleaned data

```{r}
data_location <- here::here("data","processed_data","processeddata.rds")
data<- readRDS(data_location)
```

**Reminder**: Outcome of interest is **Body Temp**; Categorical outcome is **Nausea**; Predictor= **RunnyNose**


# Module 11 Contributor: Monica Chan

## Body Temperature vs. all predictors

Below I am creating the new full model recipe for body temperature, making a path to the linear modeling, and combining them into a workflow to create a fitted object.
