---
title: "Tidy Tuesday Exercise 2"
output: 
  html_document:
    toc: FALSE
---
```{r Setup, message=FALSE, warning=FALSE}
library(tidyverse)  # data wrangling suite
library(tidymodels) # modeling suite (possibly use)
library(broom)      # Clean up outputs
library(skimr)      # Data Summary statistics
library(janitor)    # Cleaning names to make human friendly
library(here)       # try using this to find files easier
```

# Marble Racing
```{r }
# Get the data
marbles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-02/marbles.csv')
```
## Data Summary
```{r View Data}
skim(marbles)
```
```{r Original Data overview}
summary(marbles) # summary data

# Look at the unique aspects of some variables
unique(marbles$site) 
unique(marbles$team_name)
unique(marbles$marble_name)  
unique(marbles$pole)  
unique(marbles$notes)  
```
Summary of unique entries is we have 8 sites, 16 teams, and 32 marbles.

Viewed 8 unique notes. Considering removing all filled entries as they have cases that may have affected the race instead of just being marble, and site. Also include possible collisions and I want to eliminate the   

Pole signals the position/competitive situation of an entry, goes from 1-16 with some NAs. Will want to clean up separately for a different analysis potentially.

## General Cleaning
make some of the character 
```{r Variable Adjustment}
marbles%>%
  filter(notes=="NA")%>%
  mutate(marble_name=as.factor(marble_name), 
         team_name=as.factor(team_name), 
         site=as.factor(site),
         TotalTrackDistance=((track_length_m)*(number_laps))
         ) %>%
  filter(notes=="NA")%>%
  summary()
  


```


After looking at the data and key, notice that there is a notes variable. 
Operating on the assumption of no issues will likely remove any 

```{r Cleaning Script}

```



