---
title: "Tidy Tuesday Exercise 2"
output: 
  html_document:
    toc: FALSE
---
```{r Setup, message=FALSE, warning=FALSE}
library(tidyverse)  # data wrangling suite
library(tidymodels) # modeling suite (possibly use)
library(broom)      # Clean up outputs
library(skimr)      # Data Summary statistics
library(janitor)    # Cleaning names to make human friendly
library(lubridate)
library(here)       # try using this to find files easier
```

# Marble Racing
```{r }
# Get the data
marbles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-02/marbles.csv')
```
## Data Summary
```{r View Data}
skim(marbles)
```
```{r Original Data overview}
summary(marbles) # summary data

# Look at the unique aspects of some variables
unique(marbles$site) 
unique(marbles$team_name)
unique(marbles$marble_name)  
unique(marbles$pole)  
unique(marbles$notes)  
```
Summary of unique entries is we have 8 sites, 16 teams, and 32 marbles.

Viewed 8 unique notes. Considering removing all filled entries as they have cases that may have affected the race instead of just being marble, and site. Also include possible collisions and I want to eliminate the   

Pole signals the position/competitive situation of an entry, goes from 1-16 with some NAs. Will want to clean up separately for a different analysis potentially.

## General Cleaning
Make some of the character data types into factors. 
```{r Variable Adjustment}
marbles2<- marbles%>%
  mutate(
    # use lubridate to change dmy date to ymd date for ease of explorations
    date=as_date(dmy(date)),
    # chance into factors
    marble_name=as.factor(marble_name),
    team_name=as.factor(team_name),
    site=as.factor(site),
    # make new variable of total track distance traveled in race
    TotalTrackDistance=((track_length_m)*(number_laps)),
    # make new variable of Speed in m per s for total race
    Speed=((TotalTrackDistance)/(time_s))
    ) %>%
  # Filter out the notes sections with thing in col.
  filter(is.na(notes))%>%
  select(date, site, source, marble_name,team_name, time_s, pole, points, track_length_m,number_laps,avg_time_lap,TotalTrackDistance,Speed)
  
marbles2%>%
  summary()

```
### Summary of the racing sites
```{r}
marbles2%>%
  group_by(site, TotalTrackDistance, track_length_m, number_laps)%>%
  arrange(TotalTrackDistance)%>%
  summarize()

```
After looking at the data and key, notice that there is a notes variable. 
Operating on the assumption of no issues will likely remove any 

See that there are repeat info on the site of race and the speed of a single lap versus multiple laps. Will need to focus on either 1 lap or the full race. filter out the single runs

### Look at the marble speeds by the racing sites
```{r}
# speed is calculated by the total distance of a race (laps* track length)
marbles2%>%
  filter(number_laps!="1")%>% #remove the single laps
  ggplot(aes(x=site, y=Speed))+
  geom_boxplot(aes())+
  geom_jitter(aes(color=team_name))+
  labs(title = "Average speed of marbles by site",
       y="Speed (m/s)",
       x= "Racing Tracks")
```
### Points earned by marble
```{r}
speedpoints<-marbles2%>%
  filter(points!="NA")%>%
# see the team and top marble stats
  group_by(team_name, marble_name)%>%
  summarise(
    n=n(),
    "Points"=mean(points),
    "Points SD"=sd(points),
    "m/s"=mean(Speed),
    "m/s SD"=sd(Speed)
  )%>%
  arrange(desc(Points, Speed))%>%
#clarify in naming
    rename("Team Name"=team_name,
         "Marble Name"= marble_name,
         "Points Earned"= Points)
print(speedpoints)
```
Top point getters are from Savage Speeders.

Let's graphically see this:

### Plotting the Marble points earned by team w/ insights into  contributing marbles
```{r}
speedpoints%>%
  # reorder the teams by total points earned
  ggplot(aes(`Points Earned`,y=reorder(`Team Name`,`Points Earned`)))+
  #add the marble names to show what type of contribution they had per total points earned in the season
  geom_col(aes(fill=`Marble Name`))+
  # need to rename the legend
  guides(fill=guide_legend(title="Marbles"))+
  labs(title= "Marbula Marble Average Points Earned per site season",
       x= "Points Earned",
       y= "Team",
       caption = " *Two Marbles earned 0 points (Vespa of the Hornets & Mary of Team Primary)."
       )
```

Still curious about the sites, let's see how many points we are earning per race site
```{r}
marbles2%>%
  filter(number_laps!="1")%>% #remove the single laps
  # reorder the teams by total points earned
  ggplot(aes(points,y=reorder(team_name,points)))+
  #add the marble names to show what type of contribution they had per total points earned in the season
  geom_col(aes(fill=marble_name))+
  # Separate each site to see the individual points for each site
  facet_wrap(.~site)+
  labs(title= "Marbula Marble points earned at each race site",
       x= "Points Earned",
       y= NULL,
       caption = " *Two Marbles earned 0 points (Vespa of the Hornets & Mary of Team Primary)."
       )+
  # need to rename the legend
  guides(fill=guide_legend(title="Marbles"))+
  # Try to adjust the crowded Y axis with teams
  theme(axis.text.y=element_text(angle=0,size = rel(0.5), 
                                 margin = margin(-0.5, unit = "cm"),
                                 vjust =0))
```

## Additional analysis

### TBD
```{r Cleaning Script}

```



